Anti-Plagiarism System for Exam Monitoring

### 2025-06-13 Abstract.tex
feat: checking the basic requirements

\begin{abstract}
Academic integrity represents a fundamental challenge in modern education systems, with plagiarism rates increasing globally across all educational levels. Traditional exam monitoring approaches rely primarily on screen surveillance and human oversight, creating significant vulnerabilities in detecting sophisticated cheating behaviors during online and remote examinations.

This research develops a comprehensive anti-plagiarism monitoring system using advanced machine learning and computer vision technologies to address these limitations. The system integrates sophisticated gaze tracking algorithms with YOLO-based object detection models through a modular software architecture that combines facial landmark detection with real-time behavioral analysis. Custom algorithms continuously analyze candidate gaze patterns while convolutional neural networks detect unauthorized objects and suspicious materials, processing video streams efficiently using optimized OpenCV libraries.

Experimental testing demonstrates highly effective operation on standard CPU-based systems, successfully identifying suspicious gaze patterns, unauthorized devices, and anomalous behaviors while maintaining consistent real-time performance. The system achieved robust detection capabilities during controlled testing scenarios. This research provides a comprehensive open-source solution for educational institutions, enabling widespread deployment using standard computing hardware without requiring specialized equipment.
\end{abstract}

@@

Academic integrity represents a cornerstone of quality education, with educational institutions worldwide facing unprecedented challenges in maintaining fair assessment practices. The proliferation of digital learning environments has amplified concerns about academic misconduct, with studies indicating that plagiarism rates have increased significantly across all educational levels\cite{zimba2021plagiarism}. Traditional examination monitoring approaches, while foundational, present significant limitations in detecting sophisticated cheating behaviors, particularly in remote and hybrid learning contexts.

The context of plagiarism in educational settings has evolved dramatically with technological advancement. Research demonstrates that academic dishonesty affects educational quality, credibility, and institutional reputation\cite{pelican2021plagiat}. Current statistics reveal alarming trends, with some regions reporting plagiarism detection rates exceeding 26\% of submitted academic works, nearly double the global average. This widespread issue necessitates immediate technological intervention to preserve educational standards and ensure equitable assessment environments.

The problem this research addresses lies in the limitations of existing monitoring solutions, which rely heavily on manual supervision and screen-based surveillance. Traditional approaches fail to detect physical cheating behaviors, such as unauthorized device usage or suspicious gaze patterns, creating vulnerabilities in examination integrity. While machine learning solutions exist for plagiarism detection\cite{nazari2019detection}, most require extensive computational resources including GPU acceleration, making them inaccessible to many educational institutions with limited hardware infrastructure.

@@

You Only Look Once (YOLO) architectures have revolutionized real-time object detection applications. YOLO algorithms process entire images in single forward passes, enabling rapid detection of multiple object categories simultaneously\cite{wang2022object}. The latest YOLOv8 implementations demonstrate superior performance in detecting small objects and maintaining accuracy across diverse environmental conditions\cite{v7labs2023yolo}.

Convolutional Neural Networks (CNNs) provide the underlying architecture for modern object detection systems\cite{goodfellow2016deep}. These networks excel at feature extraction and pattern recognition, enabling identification of unauthorized devices such as mobile phones and smartwatches in examination environments. The hierarchical feature learning capabilities of CNNs make them particularly effective for distinguishing between similar object categories\cite{pytorch}.

@@

WebRTC protocols enable real-time multimedia communication in educational applications, providing low-latency video transmission capabilities essential for monitoring systems\cite{sredojev2015webrtc}. These technologies facilitate immediate response to detected violations while maintaining system responsiveness during extended examination periods.

@@

The facial detection module serves as the primary behavioral analysis component, utilizing dlib facial landmark detection to identify 68 key facial points. These landmarks enable precise isolation of eye regions for subsequent pupil tracking and gaze direction estimation. The module processes grayscale images to reduce computational overhead while maintaining detection accuracy.

@@

The gaze analysis subsystem implements sophisticated algorithms for determining viewing direction through pupil position analysis. The system calculates horizontal and vertical ratios based on pupil coordinates relative to eye boundaries, incorporating head orientation compensation to distinguish between natural movements and suspicious behavior patterns.

The horizontal ratio calculation is essential for determining if the candidate is looking left or right. This algorithm calculates the ratio based on pupil positions and includes adjustment components for head orientation:

@@

The object detection component utilizes dual YOLOv8 models specialized for identifying unauthorized devices. Separate models for mobile phones and smartwatches enable optimized detection thresholds and reduced false positive rates. The system processes video frames at intervals to balance detection accuracy with computational efficiency.

Real-time processing optimization includes frame skipping strategies, where object detection occurs on every 20th frame while maintaining continuous facial analysis. This approach significantly reduces computational load while preserving system responsiveness for immediate violation detection.

\subsection{Integration and Communication}

Component integration follows a publisher-subscriber pattern, enabling asynchronous communication between detection modules and the central violation monitoring system. This architecture ensures system responsiveness while accommodating varying processing times across different detection algorithms.

The violation monitoring module aggregates detection results from multiple sources, applying temporal filtering to reduce false positives caused by brief, natural movements. Alert generation includes confidence scoring and contextual information to assist human supervisors in validation decisions.

\subsection{Real-time Optimization Strategies}

Performance optimization represents a critical aspect of the system design, enabling operation on standard CPU-based hardware. The implementation employs several sophisticated optimization techniques:

\textbf{Image Preprocessing Pipeline:} Gaussian filtering with 7×7 kernels reduces noise in eye region analysis while maintaining pupil detection accuracy. The standard deviation is automatically calculated using:

\begin{equation}
\sigma = 0.3 \times \left(\frac{kernel\_size - 1}{2} - 1\right) + 0.8
\end{equation}

For 7×7 kernels: $\sigma = 0.3 \times (3-1) + 0.8 = 1.4$

\section{System Architecture Diagrams and Analysis}

\subsection{Activity Diagram and Operational Flow}

@@

\begin{figure}[H]
    \centering
    \begin{tikzpicture}[
        scale=0.7,
        transform shape,
        node distance=1.0cm,
        start/.style={circle, fill=black, minimum width=0.5cm},
        end/.style={circle, draw=black, thick, fill=white, minimum width=0.5cm},
        process/.style={rectangle, draw=black, fill=blue!30, minimum width=2.8cm, 
                        minimum height=0.8cm, text centered, rounded corners},
        decision/.style={diamond, draw=black, fill=yellow!20, text centered, 
                        minimum width=1.8cm, minimum height=0.7cm, inner sep=0pt},
        arrow/.style={thick, ->, >=stealth}
    ]
        \node[start] (initpoint) {};
        \node[process, below of=initpoint] (start) {Start Monitoring};
        \node[process, below of=start, yshift=-0.1cm] (init) {Initialize Camera};

@@

\subsection{Sequence Diagram Analysis}

@@

\begin{figure}[H]
    \centering
    \begin{tikzpicture}[
        scale=0.6,
        transform shape,
        participant/.style={rectangle, draw=black, fill=blue!30, text centered, minimum width=2.0cm, minimum height=0.6cm, font=\sffamily\small},
        activation/.style={rectangle, draw=black, fill=gray!30, minimum width=0.2cm},
        lifeline/.style={dashed},
        message/.style={-{Latex}, thick},
        return/.style={-{Latex[length=2mm]}, dashed, thick},
    ]
        \node[participant] (gui) at (0,0) {GUI};
        \node[participant] (sistem) at (2.5,0) {System};
        \node[participant] (video) at (5,0) {Video};
        \node[participant] (face) at (7.5,0) {Face};
        \node[participant] (object) at (10,0) {Object};
        \node[participant] (violation) at (12.5,0) {Monitor};

@@

\subsection{Advanced Implementation Details}

\textbf{Dual YOLOv8 Architecture Implementation:} The object detection framework employs sophisticated model specialization, with separate confidence thresholds optimized through empirical testing\cite{ultralytics}. The mobile phone detection model operates at 0.55 confidence, balancing sensitivity with false positive reduction. The smartwatch model utilizes 0.4 confidence, accommodating smaller detection targets while maintaining accuracy through secondary validation based on dimensional characteristics.

@@

\textbf{Facial Landmark Integration and Gaze Analysis:} The system leverages dlib's 68-point facial landmark detection for precise eye region isolation\cite{el2023drowsiness}. Points 36-47 define eye contours, enabling accurate pupil localization through polygonal masking and Gaussian filtering. The minimum intensity point detection method proves robust across diverse facial orientations and lighting conditions.

@@

\textbf{Violation Monitoring and Temporal Correlation:} The ViolationMonitor class implements sophisticated temporal filtering, maintaining minimum intervals between similar alerts to prevent notification flooding\cite{pythondatetime}. Alert generation includes confidence scoring and contextual information, enabling supervisors to prioritize responses based on violation severity and frequency.

@@

\textbf{Performance Metrics Collection:} Assessment focused on detection accuracy rates for both gaze direction and object identification, processing latency measurements under sustained operation, comprehensive resource utilization monitoring including CPU, memory, and storage, and detailed false positive and false negative analysis across different scenarios.

@@

\textbf{System Resource Utilization Analysis:}
- CPU utilization peaked at 60\% during intensive processing periods
- Memory consumption stabilized at 500MB average usage during extended sessions\cite{harris2020array}
- Frame processing rate sustained 15-20 FPS throughout monitoring periods
- Storage efficiency achieved 1GB per hour of recorded examination session

@@

Compared to research prototypes requiring high-end GPU resources, this implementation successfully demonstrates that effective anti-plagiarism monitoring can operate on standard institutional hardware. This accessibility factor represents a crucial advancement for educational institutions with limited technical infrastructure.

@@

@@

\bibitem{sredojev2015webrtc}
A. Sredojev et al., "WebRTC technology overview and signaling solution design and implementation," \textit{Computer Communications}, vol. 70, pp. 89-101, 2015.

\bibitem{pytorch}
A. Paszke et al., "PyTorch: An imperative style, high-performance deep learning library," \textit{Advances in Neural Information Processing Systems}, vol. 32, pp. 8024-8035, 2019.

>>>>

\begin{abstract}
Academic integrity represents a fundamental challenge in modern education systems, with plagiarism rates increasing globally across all educational levels. Traditional exam monitoring approaches rely primarily on screen surveillance and human oversight, creating significant vulnerabilities in detecting sophisticated cheating behaviors during online and remote examinations.
This research develops a comprehensive anti-plagiarism monitoring system using advanced machine learning and computer vision technologies to address these limitations. The system integrates sophisticated gaze tracking algorithms with YOLO-based object detection models through a modular software architecture that combines facial landmark detection with real-time behavioral analysis. Custom algorithms continuously analyze candidate gaze patterns while convolutional neural networks detect unauthorized objects and suspicious materials, processing video streams efficiently using optimized OpenCV libraries.
Experimental testing demonstrates highly effective operation on standard CPU-based systems, successfully identifying suspicious gaze patterns, unauthorized devices, and anomalous behaviors while maintaining consistent real-time performance. The system achieved robust detection capabilities during controlled testing scenarios. This research provides a comprehensive open-source solution for educational institutions, enabling widespread deployment using standard computing hardware without requiring specialized equipment.
\end{abstract}

@@

Academic integrity represents a cornerstone of quality education, with educational institutions worldwide facing unprecedented challenges in maintaining fair assessment practices. The proliferation of digital learning environments has amplified concerns about academic misconduct, with studies indicating that plagiarism rates have increased significantly across all educational levels\cite{zimba2021plagiarism}. Traditional examination monitoring approaches, while foundational, present significant limitations in detecting sophisticated cheating behaviors, particularly in remote and hybrid learning contexts\cite{nazari2019detection}.

The context of plagiarism in educational settings has evolved dramatically with technological advancement. Research demonstrates that academic dishonesty affects educational quality, credibility, and institutional reputation\cite{pelican2021plagiat}. Current statistics reveal alarming trends, with some regions reporting plagiarism detection rates exceeding 26\% of submitted academic works, nearly double the global average\cite{brainard2018massive}. This widespread issue necessitates immediate technological intervention to preserve educational standards and ensure equitable assessment environments.

The problem this research addresses lies in the limitations of existing monitoring solutions, which rely heavily on manual supervision and screen-based surveillance. Traditional approaches fail to detect physical cheating behaviors, such as unauthorized device usage or suspicious gaze patterns, creating vulnerabilities in examination integrity\cite{dilini2021cheating}. While machine learning solutions exist for plagiarism detection\cite{russell2020artificial}, most require extensive computational resources including GPU acceleration, making them inaccessible to many educational institutions with limited hardware infrastructure.

Recent advances in computer vision and artificial intelligence have opened new possibilities for automated monitoring systems. Studies have demonstrated the effectiveness of facial detection algorithms in real-time applications\cite{hasan2021face}, while research in eye-tracking technology has shown promising results for behavioral analysis in educational contexts\cite{el2023drowsiness}. Furthermore, object detection technologies, particularly YOLO architectures, have revolutionized real-time identification capabilities\cite{wang2022object}, making them suitable for detecting unauthorized devices in examination environments.

Educational technology integration has become increasingly sophisticated, with modern systems incorporating artificial intelligence to enhance assessment integrity\cite{honorlock2023detecting}. However, most existing commercial solutions require significant computational resources or rely on external cloud processing, limiting their accessibility to institutions with constrained infrastructure\cite{proctoru}. This technological gap highlights the need for efficient, locally-processed solutions that can operate on standard hardware while maintaining high detection accuracy.

@@

You Only Look Once (YOLO) architectures have revolutionized real-time object detection applications. YOLO algorithms process entire images in single forward passes, enabling rapid detection of multiple object categories simultaneously\cite{wang2022object}. The latest YOLOv8 implementations demonstrate superior performance in detecting small objects and maintaining accuracy across diverse environmental conditions\cite{v7labs2023yolo}.
Convolutional Neural Networks (CNNs) provide the underlying architecture for modern object detection systems\cite{goodfellow2016deep}. These networks excel at feature extraction and pattern recognition, enabling identification of unauthorized devices such as mobile phones and smartwatches in examination environments. The hierarchical feature learning capabilities of CNNs make them particularly effective for distinguishing between similar object categories\cite{paszke2019pytorch}.

@@

@@

The facial detection module serves as the primary behavioral analysis component, responsible for identifying human faces within the video stream and extracting facial landmarks necessary for subsequent analysis. The gaze analysis module processes facial landmark data to determine the direction of the candidate's attention, distinguishing between normal examination behavior and potentially suspicious activities.

The object detection framework operates independently to identify unauthorized devices within the examination environment. The violation monitoring system aggregates information from both behavioral and object detection modules to determine when examination rules have been violated. Finally, the report generation module creates comprehensive documentation of the monitoring session.

@@

The gaze analysis subsystem represents the core behavioral monitoring component of the architecture. This module processes facial landmark data to determine viewing direction through geometric analysis of eye positioning relative to facial structure. The subsystem incorporates compensation mechanisms for natural head movements to distinguish between intentional gaze direction changes and normal physiological behavior.

The module architecture supports real-time analysis while maintaining computational efficiency. Head orientation compensation ensures accurate detection across varying candidate positions and natural movement patterns. The subsystem provides directional classification capabilities for left, right, center, and downward gaze orientations.

@@

The object detection component utilizes specialized recognition models to identify unauthorized devices within the examination environment. The framework employs separate detection pathways for different device categories, enabling optimized recognition thresholds and reduced classification errors.

The architecture supports real-time processing through intelligent frame selection strategies that balance detection accuracy with computational efficiency. The framework processes video streams at optimized intervals while maintaining continuous monitoring capabilities for immediate violation detection.

\subsection{Integration and Communication Architecture}

The system architecture implements a publisher-subscriber communication pattern that enables asynchronous coordination between detection modules and the central violation monitoring system. This architectural approach ensures system responsiveness while accommodating varying processing requirements across different detection algorithms.

The violation monitoring module serves as the central coordination point, aggregating detection results from multiple independent sources. The module applies temporal analysis to reduce false positive detections caused by brief, natural movements while maintaining sensitivity to genuine violations.

\subsection{System Architecture Diagrams and Operational Analysis}

To illustrate the system structure and operation, two complementary UML diagrams have been developed: the activity diagram and sequence diagram. These representations offer different perspectives on the architecture, from operational flow to temporal component interactions.

\subsubsection{Activity Diagram and Operational Flow}

@@

\begin{figure}[H]
    \centering
    \begin{tikzpicture}[
        scale=0.5,
        transform shape,
        node distance=1.0cm,
        start/.style={circle, fill=black, minimum width=0.5cm},
        end/.style={circle, draw=black, thick, fill=white, minimum width=0.5cm},
        process/.style={rectangle, draw=black, fill=blue!30, minimum width=2.8cm, 
                        minimum height=0.8cm, text centered, rounded corners},
        decision/.style={diamond, draw=black, fill=yellow!20, text centered, 
                        minimum width=1.8cm, minimum height=0.7cm, inner sep=0pt},
        arrow/.style={thick, ->, >=stealth}
    ]
        \node[start] (initpoint) {};
        \node[process, below of=initpoint] (start) {Start Monitoring};
        \node[process, below of=start, yshift=-0.1cm] (init) {Initialize Camera};

@@

\subsubsection{Sequence Diagram Analysis}

@@

\begin{figure}[H]
    \centering
    \begin{tikzpicture}[
        scale=0.5,
        transform shape,
        participant/.style={rectangle, draw=black, fill=blue!30, text centered, minimum width=2.0cm, minimum height=0.6cm, font=\sffamily\small},
        activation/.style={rectangle, draw=black, fill=gray!30, minimum width=0.2cm},
        lifeline/.style={dashed},
        message/.style={-{Latex}, thick},
        return/.style={-{Latex[length=2mm]}, dashed, thick},
    ]
        \node[participant] (gui) at (0,0) {GUI};
        \node[participant] (sistem) at (2.5,0) {System};
        \node[participant] (video) at (5,0) {Video};
        \node[participant] (face) at (7.5,0) {Face};
        \node[participant] (object) at (10,0) {Object};
        \node[participant] (violation) at (12.5,0) {Monitor};

@@

\section{Implementation}

\subsection{Technology Stack and Platform Compatibility}

The implementation leverages Python as the primary development language, chosen for its extensive computer vision library ecosystem and rapid prototyping capabilities. The system demonstrates excellent cross-platform compatibility, operating efficiently on both Windows and Linux (Ubuntu) environments without modification to the core functionality.

\textbf{Windows Platform Performance:} On Windows systems, the application benefits from enhanced performance through CUDA acceleration when NVIDIA GPUs are available. This GPU support provides approximately 40\% improvement in processing speed for object detection tasks. The Windows implementation takes advantage of optimized driver management and CUDA toolkit integration, allowing the system to leverage hardware acceleration for both YOLOv8 model inference and OpenCV operations.

\textbf{Linux Platform Compatibility:} The system operates effectively on Linux distributions, particularly Ubuntu, maintaining full functionality through CPU-based processing. While GPU acceleration via CUDA is not yet fully implemented on Linux due to differences in driver management and CUDA installation procedures, the optimized algorithms ensure acceptable performance levels for real-time monitoring requirements. Linux deployment benefits from the platform's stability and resource management capabilities, making it suitable for institutional server environments.

\subsection{Advanced Gaze Analysis Implementation}

The gaze analysis implementation employs sophisticated mathematical algorithms for determining viewing direction through pupil position analysis. The system calculates horizontal and vertical ratios based on pupil coordinates relative to eye boundaries, incorporating head orientation compensation to distinguish between natural movements and suspicious behavior patterns.

The horizontal ratio calculation determines if the candidate is looking left or right. This algorithm processes pupil positions and includes adjustment components for head orientation:

\subsection{Dual YOLOv8 Architecture Implementation}

The object detection framework employs sophisticated model specialization, with separate confidence thresholds optimized through empirical testing\cite{ultralytics}. The mobile phone detection model operates at 0.55 confidence, balancing sensitivity with false positive reduction. The smartwatch model utilizes 0.4 confidence, accommodating smaller detection targets while maintaining accuracy through secondary validation based on dimensional characteristics.

@@

\subsection{Core Library Integration and Optimization}

\textbf{OpenCV Implementation:} The OpenCV library (\texttt{Open Source Computer Vision Library}) serves as the foundation for all image processing operations\cite{hasan2021face}. Video capture is achieved through the \texttt{VideoCapture} function, which interfaces directly with webcam hardware. The \texttt{cvtColor} function converts color images to grayscale, reducing computational complexity by processing single-channel data instead of three-channel RGB, significantly improving facial detection performance.

Gaussian filtering implementation utilizes \texttt{cv2.GaussianBlur} with 7×7 kernels for noise reduction in eye region analysis. The standard deviation is automatically calculated using the formula 
\begin{equation}
\sigma = 0.3 \times \left(\frac{kernel\_size - 1}{2} - 1\right) + 0.8
\end{equation}resulting in $\sigma = 1.4$ for 7×7 kernels. This configuration provides optimal balance between noise reduction and edge preservation for pupil detection.

The \texttt{cv2.minMaxLoc} function enables precise pupil localization by identifying minimum intensity points within isolated eye regions. This approach proves robust across varying lighting conditions when combined with Gaussian pre-filtering. Visual feedback is implemented through \texttt{cv2.circle} for pupil highlighting and \texttt{cv2.rectangle} for object detection visualization.

\textbf{Dlib Facial Landmark Detection:} The \texttt{dlib.get\_frontal\_face\_detector} initializes HOG-based face detection, optimized for real-time performance. The \texttt{dlib.shape\_predictor} loads pre-trained models for 68-point facial landmark extraction. Points 36-47 specifically define eye contours, enabling precise eye region isolation for subsequent gaze analysis. The \texttt{shape.part} function provides coordinate access for each landmark, facilitating geometric calculations for head orientation and gaze direction estimation.

\textbf{PyTorch and YOLOv8 Integration:} Model loading utilizes \texttt{torch.load} for pre-trained YOLOv8 weights, with separate models optimized for mobile phones and smartwatches. The \texttt{model.predict} method processes video frames with configurable confidence thresholds (0.55 for phones, 0.4 for watches). Frame processing optimization reduces computational load by analyzing every 20th frame while maintaining continuous facial monitoring.

\subsection{Real-time Optimization Implementation}

Performance optimization represents a critical aspect of the implementation, enabling operation on standard CPU-based hardware. The system employs several sophisticated optimization techniques:

\subsection{Facial Landmark Integration and Pupil Detection}

The system leverages dlib's 68-point facial landmark detection for precise eye region isolation\cite{el2023drowsiness}. Points 36-47 define eye contours, enabling accurate pupil localization through polygonal masking and Gaussian filtering. The minimum intensity point detection method proves robust across diverse facial orientations and lighting conditions.

@@

\subsection{Violation Monitoring and Temporal Correlation}

The ViolationMonitor class implements sophisticated temporal filtering, maintaining minimum intervals between similar alerts to prevent notification flooding\cite{pythondatetime}. Alert generation includes confidence scoring and contextual information, enabling supervisors to prioritize responses based on violation severity and frequency.

@@

@@

@@

@@

\bibitem{paszke2019pytorch}
A. Paszke et al., "PyTorch: An imperative style, high-performance deep learning library," \textit{Advances in Neural Information Processing Systems}, vol. 32, pp. 8024-8035, 2019.

@@
